---
title: "Homework 3"
author: "Zhenyu Zhou"
date: "2022-10-14"
output: github_document

---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

### Problem 1

#### Read in the data

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

#### Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
instacart %>% 
  count(aisle) %>% 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))

```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```

### Problem 2

**Load and tidy Data**

```{r}
accel_df = 
  read.csv("./data/accel_data.csv") %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_to = "minute",
    values_to = "activity_counts",
    names_prefix = "activity_"
  ) %>% 
  mutate(
    minute = as.numeric(minute),
    day = factor(day, 
                 levels = c("Monday","Tuesday","Wednesday","Thursday","Friday","Saturday","Sunday")),
    weekday_weekend = if_else((day == "Saturday" | day == "Sunday"),"Weekend","Weekday"),
    weekday_weekend = factor(weekday_weekend, 
                             levels = c("Weekday", "Weekend"))
  ) %>% 
  relocate(week,day_id, day,weekday_weekend)

```
**Describe the resulting dataset:**

The accel_df dataset contains `r nrow(accel_df)` observations of activity (accelerometer data), and there are `r ncol(accel_df)` variables, which are `r names(accel_df)` that describing the information of data. 

`r names(accel_df)[5]`: the number of minute of a 24-hour day starting at midnight.

`r names(accel_df)[6]`: the activity count of the male collected by accelerometer.

**Traditional analyses of accelerometer data focus on the total activity over the day. Using your tidied dataset, aggregate across minutes to create a total activity variable for each day, and create a table showing these totals. Are any trends apparent?**

```{r}
total_activity_data = accel_df %>% 
  group_by(week, day) %>%
  summarize(activity_total = sum(activity_counts),
            .groups = "drop") %>%
  pivot_wider(names_from = day,
              values_from = activity_total)
knitr::kable(total_activity_data)
```

The trend in total daily activity counts changed significantly over time over five weeks. The weekend accelerometer monitoring data has a tendency to be much lower than the weekday data. And, the trend in total daily activity counts changed significantly over time over five weeks.

**Make a single-panel plot that shows the 24-hour activity time courses for each day and use color to indicate day of the week.**

```{r fig.width = 10, fig.height = 9}
accel_df %>% 
  ggplot(aes(x = minute, 
             y = activity_counts, 
             color = day)) +
  geom_line(alpha = .4) +
  scale_x_continuous(
    breaks = c(0, 180, 360, 540, 720, 900, 1080, 1260, 1440),
    labels = c("0:00", "3:00", "6:00", "9:00","12:00","15:00","18:00","21:00","24:00"),
    limits = c(0,1440)) +
  labs(title = "24-hour activity time courses per day",
       x = "Time",
       y = "Activity accelerometer data") +
  theme(plot.title = element_text(hjust = 0.5))
```

From the graph, we can calculate that this male's activity count increased around 11am to 12pm and 7pm to 10pm.

### Problem 3

**Clean and load data from ny_noaa.**

```{r}
library(p8105.datasets)
data("ny_noaa")
```

This dataset provides weather data from weather stations in New York State from 1981 to 2010. This dataset has `r nrow(ny_noaa)` rows and `r ncol(ny_noaa)` columns. 

This dataset contains variables: `r names(ny_noaa)`. 

`id`: weather station ID, `date`: date of observation, `prcp`: precipitation, `snow`: snowfall, `snwd`: snow depth, `tmax`: maximum temperature, and `tmin`: minimum temperature. 

The dataset contains a lot of missing data, especially for `tmax` and `tmin`.

**Clean data, create variables and wrangle data**

```{r}
noaa_clean =   
  ny_noaa %>% 
  janitor::clean_names() %>% 
  separate(date, into = c("year", "month", "day", sep = "-")) %>% 
  mutate(
    month = month.name[as.integer(month)],
    prcp = prcp * 0.1,
    tmax = as.integer(tmax) * 0.1,
    tmin = as.integer(tmin) * 0.1
  )
```

```{r}
noaa_clean %>% 
  count(snow) %>% 
  arrange(desc(n))
```

The most commonly observed value is 0, since most of time does not snow.

**Make a two-panel plot showing the average max temperature in January and in July in each station across years. Is there any observable / interpretable structure? Any outliers?**

```{r}
noaa_clean %>% 
  filter(month %in% c("January", "July")) %>% 
  group_by(month, id, year) %>% 
  summarise(mean_tmax = mean(tmax, na.rm = TRUE)
  ) %>% 
  ggplot(aes(x = as.integer(year), y = mean_tmax, group = id,color = month)) +
  geom_line(alpha = 0.3) +
  facet_grid(. ~ month, scales = "free") +
  labs(
    x = "Year",
    y = "Average Max Temperature (°C)",
    title = "Average daily maximum temperature in New York State"
  ) +
  theme(legend.position = "none")
```

The average maximum temperature in January is between -10-10°C. The average maximum temperature in July is between 20-32°C. The change in tmax in January is greater than that in July. The most outlier in January is around -14°C and the most outlier in July is around 13°C.

**Make a two-panel plot showing (i) tmax vs tmin for the full dataset (note that a scatterplot may not be the best option); and (ii) make a plot showing the distribution of snowfall values greater than 0 and less than 100 separately by year.**

```{r}
p_tmax_tmin =   
  noaa_clean %>% 
  ggplot(aes(x = tmin, y = tmax)) +
  geom_hex() +
  labs(
    title = "Tmax vs Tmin temperature",
    x = "Min temperature (°C)",
    y = "Max temperature (°C)"
  ) +
  theme(legend.position = "left")

p_snowfall =
  noaa_clean %>% 
  filter(snow > 0 & snow < 100) %>% 
  group_by(year) %>%
  ggplot(aes(x = snow, y = year)) +
  geom_density_ridges(alpha = 0.5) +
  labs(
    title = "Snowfall by year", 
    x = "Snowfall (mm)", 
    y = "Year") +
  scale_x_discrete(
    breaks = c(0,25,50,75,100)) +
  scale_y_discrete(
    position = "left") +
  viridis::scale_fill_viridis() 

p_tmax_tmin + p_snowfall
```

